---
title: "Assignment-7"
author: "ItsmeHJB"
date: "18/11/2021"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(palmerpenguins)

data(package = 'palmerpenguins')
```

# Assignment 7

## Student's t-confidence intervals

Compute confidence intervals of flippers using pop mean of flipper length utilising Student's t-distribution. This requires some data setup using palmerpenguins.

```{r data creation}
# Select adelie penguin's flipper length
adelie_flippers <- penguins %>%
  filter(species=="Adelie") %>%
  select(flipper_length_mm)

# Convert data frame to vector
adelie_flippers <- adelie_flippers$flipper_length_mm
# Remove na data
adelie_flippers <- adelie_flippers[!is.na(adelie_flippers)]
```


```{r t-confidence intervals}
# Setup alpha and other required vars
alpha<-0.05
sample_size<-length(adelie_flippers)
sample_mean<-mean(adelie_flippers)
sample_sd<-sd(adelie_flippers)

# Calc t value and conf interval
t<-qt(1-alpha/2,df=sample_size-1)
confidence_interval_l<-sample_mean-t*sample_sd/sqrt(sample_size)
confidence_interval_u<-sample_mean+t*sample_sd/sqrt(sample_size)
confidence_interval<-c(confidence_interval_l,confidence_interval_u)
confidence_interval
```

Doing the same again with the hawks dataset, specifically the red-tailed hawks and a 99% confidence level.
```{r hawks t-test}
library(Stat2Data)
data("Hawks")

# Select red-tailed hawks weights
rt_weight <- Hawks %>%
  filter(Species=="RT") %>%
  select(Weight)

# Convert data frame to vector
rt_weight <- rt_weight$Weight
# Remove na data
rt_weight <- rt_weight[!is.na(rt_weight)]

# Setup alpha and other required vars
# 99%-level confidence intervals
alpha<-0.01
sample_size<-length(rt_weight)
sample_mean<-mean(rt_weight)
sample_sd<-sd(rt_weight)

# Calc t value and conf interval
t<-qt(1-alpha/2,df=sample_size-1)
confidence_interval_l<-sample_mean-t*sample_sd/sqrt(sample_size)
confidence_interval_u<-sample_mean+t*sample_sd/sqrt(sample_size)
rt_confidence_interval<-c(confidence_interval_l,confidence_interval_u)
rt_confidence_interval
```

This is all assuming the data is Gaussian by nature.

## One sample t-test

Load bill length of adelie penguins. Carry out one sample t-test on data using t.test function.

```{r t-test}
# Select adelie penguin's bill length
adelie_bills <- penguins %>%
  filter(species=="Adelie") %>%
  select(bill_length_mm)

# Convert data frame to vector
adelie_bills <- adelie_bills$bill_length_mm
# Remove na data
adelie_bills <- adelie_bills[!is.na(adelie_bills)]

t.test(x=adelie_bills, mu=40, conf.level = 0.99)
```

As the mean falls outside of the 99% confidence interval, we reject the null hypothesis that mu = 40, and accept the alternative.

We are assuming the data follows a Gaussian distribution

## Implementing one-sample t-test

The function takes an input of x, a Gaussian distribution of random variables in the shape of a vector, and a mean, mu, of the null hypothesis. It returns the p-value of the test, which can be compared against the t.test function's p-value.

```{r}
# Pass in vector x and value, mu0, corresponding to the value of the null hypothesis
t_test <- function(x, mu0) {
  alpha = 0.05
  sample_size <- length(x)
  sample_mean <- mean(x, na.rm = TRUE)
  sample_sd <- sd(x, na.rm = TRUE)
  
  test_stat <- (sample_mean-mu0) / (sample_sd/sqrt(sample_size))
  
  p_value = 2 * (1 - pt(abs(test_stat), df = sample_size - 1))
  
  return(p_value)
}

t_test(adelie_bills, 40)
t.test(x=adelie_bills, mu=40)
```

## Paired test

Compare the two barley yields in two fields. t.test is used on the data to see that there is a significant difference between the two variables.

```{r data loading}
library(PairedData)
data("Barley")
```

```{r}
alpha <- 0.01
diffs <- Barley %>%
  mutate(diff = Velvet - Glabron) %>%
  dplyr::select(diff)
diffs <- diffs$diff

t.test(x=Barley$Glabron, y=Barley$Velvet, paired=TRUE, conf.level = 1-alpha)
```

Afterwards, we work out the cohen's d statistic

```{r}
y_bar <- mean(diffs)
s <- sd(diffs)
effect_size <- abs(y_bar / s)
effect_size
```

As the effect_size is large, we can say they are significantly different.

We have assumed that the sample follows the Gaussian distribution.



