---
title: "Assignment-9"
author: "ItsmeHJB"
date: "01/12/2021"
output: html_document
---

```{r setup, include=FALSE}
library(Stat2Data)
```

## Basic concepts in classification

1. Classification rule - A rule which is able to map a feature vector to categorical values.

2. Learning algorithm - An algorithm which machines can use to learn to classify data.

3. Training data - Data used to train a model.

4. Feature vector - Vector containing feature x1 to xn.

5. Label - Each set of features is assigned one label.

6. Test error - The number of incorrect labels created by the model in comparison to the actual labels in the test data.

7. Train error - The number of incorrect labels created by the modes in comparison to the actual labels in the training data.

8. Train test split - Splitting the original data set into a training and test set.

9. Linear classifier - Cuts the feature space into two with a linear hyper-plane.

## Train test split

Get data frame containing weight, wing, hallux, tail and species. This should only include sharp-shinned or cooper's species.

```{r}
data("Hawks")
```


