---
title: "Assignment-5"
author: "ItsmeHJB"
date: "27/10/2021"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(ggplot2)
```

# Assignment 5
## 1 - Expectation and variance in discrete random variable

Let $\alpha,\beta \in [0,1]$ with $\alpha + \beta \le 1$
X is supported on {0, 1, 5}

We could also say: $\Omega = \{0, 1, 5\}$

$\mathbb{P}(X=1)=\alpha$ and $\mathbb{P}(X=5)=\beta$

Also,
$\mathbb{P}(X \notin \{0, 1, 5\}) = 0$

### Probability mass function

Define the prob for each value of $\Omega$
\begin{equation}
  \mathbb{P}(X) =
  \begin{cases}
    1 - \alpha - \beta & x=0 \\
    \alpha & x=1 \\
    \beta & x=5 \\
    0 & \text{otherwise}
  \end{cases}
\end{equation}

$px(S) = (1 - \alpha - \beta) + \alpha$

$px(S) = 1 - \beta$

### Expectation

$\mathbb{E}(X) = (0 \cdot 1-\alpha-\beta) + (1 \cdot\alpha) + (5\cdot\beta) $

$\mathbb{E}(X) = \alpha + 5\beta $

### Variance

$Var(X) = \mathbb{E}(X^2) - [\mathbb{E}(X)]^2$

Work out E(X)^2
$\mathbb{E}(X^2) = (0^2 \cdot 1-\alpha-\beta) + (1^2 \cdot\alpha) + (5^2\cdot\beta) $

$\mathbb{E}(X^2) = 0 + \alpha + 25\beta$

$\mathbb{E}(X^2) = \alpha + 25\beta$

$Var(X) = \alpha + 25\beta - (5\beta+\alpha)^2$

$Var(X) = \alpha + 25\beta - (25\beta^2+10\beta\alpha+\alpha^2)$

## 2 - Simulating data with uniform distribution

```{r}
set.seed(0)

n<-1000

sample_X<-data.frame(U=runif(n)) %>%
  mutate(X=case_when(
    (0<=U)&(U<0.25)~1,
    (0.25<=U)&(U<0.5)~5,
    (0.5<=U)&(U<=1)~0)) %>%
  pull(X)

```

Simulating question 1 above, we can use the above code chunk to create i.i.d. variables

In this sample, both $\alpha$ and $\beta$ equal 0.25. If the value for U falls between 0 and 0.25, it is assigned 1, between 0.25 and 0.5, it is assigned 1. And finally if it is between 0.5 and 1, it is assigned 0.

Below is a function which can take alpha, beta and n as variables.
```{r}
sample_X_015 <- function(alpha, beta, n) {
  beta_start <- alpha+beta
  
  return(data.frame(u=runif(n)) %>%
    mutate(x=case_when(
      (0<=u)&(u<alpha) ~ 1,
      (alpha<=u)&(u<beta_start) ~ 5,
      (beta_start<=u)&(u<=1) ~ 0
    )) %>%
    pull(x)
  )
}
```

Setting up the probs and running the code
```{r}
alpha <- 0.5
beta <- 0.1
n <- 10000

# Vector of probs
PS <- c(1-alpha-beta, alpha, beta)

# Vector of outcomes
S <- c(0, 1, 5)

# Run the code
vals <- sample_X_015(alpha, beta, n)

# Expectation
ES <- sum(S * PS)
ES

# Average of values
mean(vals)

# Variance
var(vals)

```

We find the output of both the expectation and average to be extremely similar here. With the average floating around 1.0

Due to the large number of experiments, the sample slowly drifts closer to the expectation. If we increased n further, it would be closer still.

Varying beta, we get
```{r}
beta_vals <- c(seq(0, 0.9, 0.01))
alpha <- 0.1
n <- 100

#temp <- data.frame(map(.x=data_x, ~sample_X_015(alpha,.x,n)))
#temp

vary_beta_sample<-data.frame(beta_vals)%>%
  mutate(rand_vars=map(.x=beta_vals,~sample_X_015(alpha, .x, n))) %>%
  mutate(sample_avg = map_dbl(.x=rand_vars,~mean(.x)))

head(vary_beta_sample)

graph <- vary_beta_sample %>%
  ggplot(aes(x=beta_vals, y=sample_avg)) + xlab("Beta")
graph + geom_line() + ylab("Sample avg")
```

## 3 - Gaussian distribution

Probability density function of the Gaussian random variable is:

$f\mu,\sigma (x) = \frac{1}{\sigma \sqrt{2\pi}} \cdot exp(-\frac{1}{2}(\frac{x-\mu}{\sigma})^2)$

The normal distribution functions all sit under the norm group:

dnorm, pnorm, qnorm and rnorm

Plotting a gaussian density func is as below:

```{r}
plot1 <- data.frame(mu = 1, sigma = 1)
plot2 <- data.frame(mu = 1, sigma = 2)
plot3 <- data.frame(mu = 1, sigma = 3)

ggplot(data.frame(x = c(-4, 6)), aes(x = x)) + 
  xlim(c(-4, 6)) + 
  stat_function(fun = dnorm, args = list(mean = 1, sd = sqrt(1)), aes(colour = "red")) + 
  stat_function(fun = dnorm, args = list(mean = 1, sd = sqrt(2)), aes(colour = "green")) +
  stat_function(fun = dnorm, args = list(mean = 1, sd = sqrt(3)), aes(colour = "blue")) +
  labs(x = "X", y = "Density", 
       title = "Gaussian or Normal distribution plots")
```

