---
title: "Assignment-6"
author: "ItsmeHJB"
date: "10/11/2021"
output: html_document
---

# Assignment 6

## Gaussian Model for Red tailed hawks

```{r setup, include=FALSE}
library(tidyverse)
library(Stat2Data)
data("Hawks")
```

Extract all red-tailed hawks, filter out Weight, Tail and Wing

```{r}
RedTailedDf <- Hawks %>%
  filter(Species=="RT") %>%
  select(Weight, Tail, Wing)
```

Calculating the mean and variance using both the formula and in-built functions produce the same answer

```{r}
mean_mu = mean(RedTailedDf$Tail)
mu = sum(RedTailedDf$Tail) / length(RedTailedDf$Tail)

var_sigma_sqrd = var(RedTailedDf$Tail)
sigma_sqrd = sum(RedTailedDf$Tail - mean_mu) ** 2 / length(RedTailedDf$Tail)

```

Simulating the MLE function vs the actual values:
```{r}
sigma <- sqrt(sigma_sqrd)  # Get standard deviation
x <- seq(mu-3*sigma, mu+3*sigma, sigma*0.01) # Generate x plots, define bounds based on 3x standard dev, and the step as 0.01 of sd
df_gaussian <- data.frame(x, Density = dnorm(x, mean=mu, sd=sigma), Source="Population") # data frame using pop density

set.seed(1) # reproducible seed
sample_size <- 100 # n
sample_data <- rnorm(sample_size, mu, sigma) # Generate data

# get mle mean and sigma
mu_mle <- mean(sample_data)
sigma_mle <- sd(sample_data)*sqrt((sample_size-1) / sample_size)

df_gaussian <- df_gaussian %>%
  rbind(data.frame(x, Density=dnorm(x, mean=mu_mle, sd=sigma_mle), Source = "MLE estimate"))  # add in mle density

# Plot it
df_gaussian %>% ggplot(aes(x=x, y=Density, color=Source)) + 
  geom_line() + 
  ylab("Density function") +
  theme_bw()
```

Doing the same again using a parametric model
```{r}
RT_tails <- RedTailedDf %>% pull(Tail)

n <- length(RT_tails)
mu_mle_tails <- mean(RT_tails, na.rm=1)
sigma_mle_tails <- sd(RT_tails)*sqrt((n-1) / n)

tails <- seq(mu_mle_tails-3*sigma_mle_tails, mu_mle_tails+3*sigma_mle_tails, sigma_mle_tails*0.001) # generate indecies

colors <- c("MLE density" = "red", "Kernel density" = "blue") # line colours
ggplot() + geom_line(data = data.frame(Tail=tails, Density=dnorm(tails, mean=mu_mle_tails, sd = sigma_mle_tails)), 
  aes(x=Tail, y=Density, color="MLE density")) + 
  geom_density(data=tibble(RT_tails), aes(x=RT_tails, color="Kernel_density")) + 
  labs(y = "Density function", color = "Estimator") + theme_bw() + 
  scale_color_manual(values=colors)
```

## Location estimators with Gaussian data

The population median of a gaussuan random variable is simply the mean

Estimate the mean squared error of the sample median as an estimator of mu

```{r}
# Setup
set.seed(0)
num_trials_per_sample_size<-100
min_sample_size<-5
max_sample_size<-1000
sample_size_inc<-5
mu_0<-1
sigma_0<-3

# Doing the mean square error of the sample median
simulation_df<-crossing(trial=seq(num_trials_per_sample_size),
                        sample_size=seq(min_sample_size,
                                       max_sample_size,sample_size_inc))%>%
  # create data frame of all pairs of sample_size and trial
  mutate(simulation=pmap(.l=list(trial,sample_size),
                          .f=~rnorm(.y,mean=mu_0,sd=sigma_0)))%>%
  # simulate sequences of Gaussian random variables
  mutate(sample_md=map_dbl(.x=simulation,.f=median))%>%
  # compute the sample medians
  group_by(sample_size)%>%
    summarise(msq_error_md=mean((sample_md-mu_0)**2))


mean_simulation_df<-crossing(trial=seq(num_trials_per_sample_size),
                        sample_size=seq(min_sample_size,
                                        max_sample_size,sample_size_inc))%>%
  # create data frame of all pairs of sample_size and trial
  mutate(simulation=pmap(.l=list(trial,sample_size),
                          .f=~rnorm(.y,mean=mu_0,sd=sigma_0)))%>%
  # simulate sequences of Gaussian random variables
  mutate(sample_mean=map_dbl(.x=simulation,.f=mean))%>%
  # compute the sample medians
  group_by(sample_size)%>%
    summarise(msq_error_mean=mean((sample_mean-mu_0)**2))

# Combine data frames
simulation_df$msq_error_mean <- mean_simulation_df$msq_error_mean

colours <- c("Mean" = "red", "Median" = "blue") # line colours
ggplot() + geom_line(data = simulation_df, 
  aes(x=sample_size, y=msq_error_md, color="Median")) + 
  geom_line(data=simulation_df, aes(x=sample_size, y=msq_error_mean, color="Mean")) + 
  labs(y = "Mean square error", x = "Sample size", color = "Estimator") + 
  theme_bw() +scale_color_manual(values=colours) + 
  xlim(0, 500) + ylim(0, 0.35)
```

## Unbiased estimation of population variance

```{r}
VMLE <- function(X) {
  
  x_bar <- mean(x)
  return(sum(x-x_bar)**2 / length(x))
}

VU <- function(X) {
  x_bar <- mean(x)
  return(sum(x-x_bar)**2 / (length(x)-1))
}

mu <- seq(1, 10)
sigma <- seq(1, 10)
sample_size = 1000

comp <- crossing(mu, sigma) %>%
  mutate(sample=map2(.x=mu, .y=sigma, ~rnorm(sample_size, mean=.x, sd=.y))) %>%
  mutate(v1=map_dbl(.x=sample, ~VMLE(.x))) %>%
  mutate(v2=map_dbl(.x=sample, ~VU(.x)))
  
comp
```

