---
title: "Assignment 3"
author: "ItsmeHJB"
date: "11/10/2021"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
```
# Part 1
## Defintions
Random experiment - A procedure that could be repeated an arbitrary number of times with a well-defined set of outcomes. E.g. flipping a coin
Event - A set of possible outcomes
Sample space - The combination of all sets

# Part 2
## Iterations
Create functions to iterate over vector using iteration
```{r}
impute_by_mean<-function(x){
  mu<-mean(x,na.rm=1) # first compute the mean of x
  impute_f<-function(z){ # coordinate-wise imputation
    if(is.na(z)){
      return(mu) # if z is na replace with mean
    }else{
      return(z) # otherwise leave in place
    }
  }
  return(map_dbl(x,impute_f)) # apply the map function to impute         across vector
}

impute_by_median<-function(x){
  mu<-median(x,na.rm=1) # first compute the mean of x
  impute_f<-function(z){ # coordinate-wise imputation
    if(is.na(z)){
      return(mu) # if z is na replace with mean
    }else{
      return(z) # otherwise leave in place
    }
  }
  return(map_dbl(x,impute_f)) # apply the map function to impute         across vector
}
```

Create 2 vectors, then combine into a data frame
A sequence of 0 to 10 in 0.1 steps
Plus a second which uses the values derived from the first
```{r}
x <- seq(0, 10, by=0.1)

to_y <- function(x) {
  return(5*x+1)
}
# Test it
to_y(0.0)
to_y(0.1)

# Map x to y
y<-map_dbl(x,to_y)

# Combine
df_xy <- data.frame(x, y)
head(df_xy, 5)
```

Create a function to return missing values when index is multiple of 5
```{r}
something_missing <- function(index, value) {
  if(index%%5 == 0) {return(NA)}
  else {return(value)}
}
```
Set up df_xy_missing the same as before
```{r}
df_xy_missing <- data.frame(x, y)
```
Start manipulating y to have missing values when index %% 5 == 0
```{r}

```

Screw this ^
#Part 3
## Elementary set theory
