---
title: "Assignment-9"
author: "ItsmeHJB"
date: "01/12/2021"
output: html_document
---

```{r setup, include=FALSE}
library(Stat2Data)
library(tidyverse)

data("Hawks")
```

## Basic concepts in classification

1. Classification rule - A rule which is able to map a feature vector to categorical values.

2. Learning algorithm - An algorithm which machines can use to learn to classify data.

3. Training data - Data used to train a model.

4. Feature vector - Vector containing feature x1 to xn.

5. Label - Each set of features is assigned one label.

6. Test error - The number of incorrect labels created by the model in comparison to the actual labels in the test data.

7. Train error - The number of incorrect labels created by the modes in comparison to the actual labels in the training data.

8. Train test split - Splitting the original data set into a training and test set.

9. Linear classifier - Cuts the feature space into two with a linear hyper-plane.

## Train test split

Get data frame containing weight, wing, hallux, tail and species. This should only include sharp-shinned or cooper's species.

```{r}
# Data pre-processing
hawks_total <- Hawks %>%
  select(Weight, Wing, Hallux, Tail, Species) %>% 
  filter(Species != "RT") %>% 
  drop_na() %>% 
  mutate(Species=as.numeric(Species=="SS"))

# 60% training data
num_total <- nrow(hawks_total)
num_train <- floor(num_total * 0.60)
num_test <- num_total - num_train

set.seed(1)
test_indi <- sample(seq(num_total), num_test)
train_indi <- setdiff(seq(num_total), test_indi)

# get train and test data
train_hawks <- hawks_total %>% 
  filter(row_number() %in% train_indi)

test_hawks <- hawks_total %>% 
  filter(row_number() %in% test_indi)

# Extract features and labels column
train_x <- train_hawks %>% select(!Species)
train_y <- train_hawks %>% pull(Species)

test_x <- test_hawks %>% select(!Species)
test_y <- test_hawks %>% pull(Species)

# Get most common label from training data
y <- round(sum(train_y) / num_train)

train_error <- mean(abs(y - train_y))
train_error
test_error <- mean(abs(y - test_y))
test_error
```

Using a dumb classifier which simply selects the most common label found in the training data, we get an test error of around 20% for the test data, and 22% for the training data. As a baseline this is interesting as it should be impossible to get worse than this using statistical methods.

From this, we can say that the data sets are relatively different sizes. Data set of type y=1, which is the Sharp-Shinned species, make up about 80% of the total data set.

## Linear discriminant analysis


